Class-based network segregation, Economic Inequality and Redistributive Preferences across societies

### 2024-03-28 paper1_jiturra.Rmd
descripcion tabla1

texreg(digclas1_VW,stars = c(0.001, 0.01, 0.05, 0.1),
       caption = "Multilevel models for Class-based network segregation and Redistributive Preferences",
       custom.note = "%stars. Models include sampling weights. Gender, age, marital status and religion are included as controls",
       caption.above = T,leading.zero = T,
       custom.coef.map = ccoef,
       symbol = "+",scalebox = 0.60,
      groups = list("Social Class (Ref.= Service Class)" = 4:5, "Household Income (Ref.= Tertile I)"= 7:9,  "Partner's Social Class (Ref.= Service Class)" = 12:14, "Homogeneity x Social Class"=15:16),
      custom.gof.rows = list("Controls"=rep("Yes",6)),include.loglik = FALSE,include.aic = FALSE,
      custom.gof.names = c(NA,NA,"Num. groups","Var: Country (Intercept)","Var: Residual"),
       float.pos = "h!",
       use.packages = F)
```

@@

```{r oldcode, eval=FALSE, include=FALSE}

# homo_V_gini_gdp_tax <- 
#   lmer(egal~ 
#          homclass3_V_gc+
#          know_total_gc+
#          socialtrust_gc+
#          dclass3res_V+dclass3spo_V+
#          female_gc+agenum_gc+age2_gc+
#          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
#          union_gc+workst_gc+
#          gini_disp +loggdppercapita + ilo_taxrev +
#          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)
# 
# 
# homo_V_gini_gdp_govsp <- 
#   lmer(egal~ 
#          homclass3_V_gc+
#          know_total_gc+
#          socialtrust_gc+
#          dclass3res_V+dclass3spo_V+
#          female_gc+agenum_gc+age2_gc+
#          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
#          union_gc+workst_gc+
#          gini_disp +loggdppercapita + ilo_govspe +
#          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)
# 
# 
# homo_V_gini_gdp_redis <- 
#   lmer(egal~ 
#          homclass3_V_gc+
#          know_total_gc+
#          socialtrust_gc+
#          dclass3res_V+dclass3spo_V+
#          female_gc+agenum_gc+age2_gc+
#          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
#          union_gc+workst_gc+
#          gini_disp +loggdppercapita + abs_red +
#          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)
# 
# homo_V_gini_gdp_controls <-
#   lmer(egal~
#          homclass3_V_gc+
#          know_total_gc+
#          socialtrust_gc+
#          dclass3res_V+dclass3spo_V+
#          female_gc+agenum_gc+age2_gc+
#          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
#          union_gc+workst_gc+
#          gini_disp +loggdppercapita + ilo_taxrev + ilo_govspe + abs_red +
#          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)
# 
# 
# homo_V_gini_gdp_wstate1 <- 
#   lmer(egal~ 
#          homclass3_V_gc+
#          know_total_gc+
#          socialtrust_gc+
#          dclass3res_V+dclass3spo_V+
#          female_gc+agenum_gc+age2_gc+
#          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
#          union_gc+workst_gc+
#          gini_disp +loggdppercapita + wstate1+
#          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

# homo_V_gini_full <- 
#   lmer(egal~ 
#          homclass3_V_gc+
#          know_total_gc+
#          socialtrust_gc+
#          dclass3res_V+dclass3spo_V+
#          female_gc+agenum_gc+age2_gc+
#          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
#          union_gc+workst_gc+
#          gini_disp +loggdppercapita + ilo_taxrev + ilo_govspe + abs_red +
#          homclass3_V_gc*
#          dclass3res_V*
#          gini_disp + 
#          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)
# 
# homo_V_gini_full1 <- 
#   lmer(egal~ 
#          homclass3_V_gc+
#          know_total_gc+
#          socialtrust_gc+
#          dclass3res_V+dclass3spo_V+
#          female_gc+agenum_gc+age2_gc+
#          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
#          union_gc+workst_gc+
#          gini_disp +loggdppercapita + wstate1+
#          homclass3_V_gc*
#          dclass3res_V*
#          gini_disp + 
#          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)
```

```{r oldcode2, eval=FALSE, fig.cap='Bivariate relationships between Income Inequality, fig.dim=c(12,5), include=FALSE, Social Class and Network homogeneity',warning=FALSE}
homo_ineq <- dfreg %>%
  # filter(country3 != "ZAF") %>%
  group_by(country2,country3,dclass3res_V) %>%
  summarise_at(c("homclass3_V_res","egal","gini_disp"), mean, na.rm = TRUE) %>% 
  as.data.frame()

homo_ineq$country <- countrycode::countrycode(sourcevar = homo_ineq$country2,origin = "iso3c",destination = "iso2c")
homo_ineq$dclass3res_V <- forcats::fct_rev(homo_ineq$dclass3res_V)
homo_ineq %>%
ggplot(aes(y=homclass3_V_res, x=gini_disp, group=dclass3res_V,fill=dclass3res_V)) +
  geom_text_repel(label=as.character(homo_ineq$country)) +
  geom_smooth(method = "lm",fullrange=TRUE, color="black",linetype = 2,size=1,se = FALSE) +
  facet_wrap(~dclass3res_V) +
  xlab("Income Inequality")+
  ylab("Network homogeneity") +
  # labs(title = "Class homogeneity and Gini (post-taxes)") +
  scale_fill_manual(values=c("#CCCCCC","#989898","#323232"))+
  scale_color_manual(values=c("#CCCCCC","#989898","#323232")) +
  theme(legend.position = "none",legend.title = element_blank())

>>>>

texreg(digclas1_VW,stars = c(0.001, 0.01, 0.05, 0.1),
       caption = paste("(\\#tab:table1)","Multilevel models for Class-based network segregation and Redistributive Preferences"),
       custom.note = "%stars. Models include sampling weights. Gender, age, marital status and religion are included as controls",
       caption.above = T,leading.zero = T,
       custom.coef.map = ccoef,
       symbol = "+",scalebox = 0.60,
      groups = list("Social Class (Ref.= Service Class)" = 4:5, "Household Income (Ref.= Tertile I)"= 7:9,  "Partner's Social Class (Ref.= Service Class)" = 12:14, "Homogeneity x Social Class"=15:16),
      custom.gof.rows = list("Controls"=rep("Yes",6)),include.loglik = FALSE,include.aic = FALSE,
      custom.gof.names = c(NA,NA,"Num. groups","Var: Country (Intercept)","Var: Residual"),
       float.pos = "h!",
       use.packages = F)
```

The multilevel analysis results are shown in Table \@ref(tab:table1). Regarding our first hypothesis ($H_1$), our central argument is that class segregation has differential consequences on redistributive preferences. In other words, the association between network homogeneity and redistributive preferences is conditional on class position.

Model 1 integrates our measure of network homogeneity. Initially, as previously noted in the macro associations, individuals with more homogeneous networks tend to exhibit stronger redistributive preferences. This association persists in Model 2, where we control for network size and social trust—two pivotal factors in establishing and consolidating interpersonal networks. However, after including social class into the analysis, the relationship between homogeneity and redistributive preferences appears to reverse, with a reduction in statistical significance (p<0.10).

Subsequently, Model 4 incorporates socioeconomic status, employment status, and union membership characteristics. As expected, socioeconomic status is negatively associated with redistributive preferences regarding household economic resources and educational credentials. In addition, we find that people outside the labor market do not differ significantly from those who are economically active in their preferences. In contrast, as expected, being unionized increases redistributive preferences compared to those who have never been union members. Then, Model 5 incorporates the social class of the respondent's partner if such information is available. Thus, it is observed that the social class of the partner follows a similar pattern to that of the respondent, where belonging to the working or intermediate class presents higher redistributive preferences.

Finally, in Model 6, we see that by incorporating the interaction term between network homogeneity and social class, we observe that the working and intermediate class interaction terms are positive and statistically significant. Figure \@ref(fig:fig4) shows increased redistributive preferences in homogeneous working-class networks and the intermediate class. For the service class, the relationship is reversed, where more homogeneous networks for this social class decrease redistributive preferences. Altogether, these results provide evidence of our hypothesis, where a higher level of network segregation is related to a greater polarization of social class interests.

@@

\pagebreak 
## Alternative models {-}

```{r tab2-ratio9010, results='asis', cache=TRUE}
set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1")
options(ggplot2.discrete.fill = set1)
options(ggplot2.discrete.colour = set1)
ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(text=ggplot2::element_text(size=15,  family="serif"))
options(scipen=999)
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal,texreg,MLMusingR,lme4)  
rm(list=ls())

dfreg <- df1 %>% dplyr::select(
  egal = egal2,
  # starts_with("class3"),
  # starts_with("digclass3"),
  # starts_with("dclass3"),
  # starts_with("homclass"),
  dclass3res_V,
  dclass3spo_V,
  homclass3_V_res,
  know_total,
  socialtrust,
  Q03pcm,
  edyears,
  female,
  agenum,
  religion,
  partner,
  union,
  workst,
  WEIGHT,
  region,
  "gini_disp"=wid_gini_disp,
  "gini_mkt",
  gv_spen,
  rel_red,
  abs_red,
  ilo_taxrev,
  ilo_govspe,
  d10d1=wid_p90p10,
  wid_p90p50,
  top10=wid_sharetop10,
  rgdpna,
  gdppercapita,
  oecd,
  country2, country, country3,
) %>%
  # filter(oecd == "OECD") %>%
  filter(country2 != "SVN") %>% 
  na.omit() %>% 
    mutate(loggdppercapita=scale(gdppercapita/1000),
         edyears2 = edyears ^ 2,
         age2 = agenum ^ 2, 
         gini_disp=scale(d10d1))

ccoef <- list(homclass3_V_gc="Class-based network homogeneity (CWC)",
              "dclass3res_VIntermediate class (III+IV)"="Intermediate Class",
              "dclass3res_VWorking Class (V+VI+VII)"="Working Class",
              "gini_disp" = "Income inequality (Ratio 90/10)",
              loggdppercapita = "GDP/capita",
              wstate2="Size of the welfare state",
              "homclass3_V_gc:dclass3res_VIntermediate class (III+IV)" = "Homogeneity*Intermediate Class",
              "homclass3_V_gc:dclass3res_VWorking Class (V+VI+VII)" = "Homogeneity*Working Class",
              "homclass3_V_gc:gini_disp"= "Homogeneity*Income Inequality",
              "homclass3_V_gc:dclass3res_VIntermediate class (III+IV):gini_disp" = "Homogeneity*Intermediate Class*Income Inequality",
              "homclass3_V_gc:dclass3res_VWorking Class (V+VI+VII):gini_disp" = "Homogeneity*Working Class *Income Inequality")

```{r tab2-ratio9050, results='asis', cache=TRUE}
set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1")
options(ggplot2.discrete.fill = set1)
options(ggplot2.discrete.colour = set1)
ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(text=ggplot2::element_text(size=15,  family="serif"))
options(scipen=999)
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal,texreg,MLMusingR,lme4)  
rm(list=ls())

dfreg <- df1 %>% dplyr::select(
  egal = egal2,
  # starts_with("class3"),
  # starts_with("digclass3"),
  # starts_with("dclass3"),
  # starts_with("homclass"),
  dclass3res_V,
  dclass3spo_V,
  homclass3_V_res,
  know_total,
  socialtrust,
  Q03pcm,
  edyears,
  female,
  agenum,
  religion,
  partner,
  union,
  workst,
  WEIGHT,
  region,
  "gini_disp"=wid_gini_disp,
  "gini_mkt",
  gv_spen,
  rel_red,
  abs_red,
  ilo_taxrev,
  ilo_govspe,
  d10d1=wid_p90p10,
  wid_p90p50,
  top10=wid_sharetop10,
  rgdpna,
  gdppercapita,
  oecd,
  country2, country, country3,
) %>%
  # filter(oecd == "OECD") %>%
  filter(country2 != "SVN") %>% 
  na.omit() %>% 
    mutate(loggdppercapita=scale(gdppercapita/1000),
         edyears2 = edyears ^ 2,
         age2 = agenum ^ 2, 
         gini_disp=scale(wid_p90p50))

ccoef <- list(homclass3_V_gc="Class-based network homogeneity (CWC)",
              "dclass3res_VIntermediate class (III+IV)"="Intermediate Class",
              "dclass3res_VWorking Class (V+VI+VII)"="Working Class",
              "gini_disp" = "Income inequality (Ratio 90/50)",
              loggdppercapita = "GDP/capita",
              wstate2="Size of the welfare state",
              "homclass3_V_gc:dclass3res_VIntermediate class (III+IV)" = "Homogeneity*Intermediate Class",
              "homclass3_V_gc:dclass3res_VWorking Class (V+VI+VII)" = "Homogeneity*Working Class",
              "homclass3_V_gc:gini_disp"= "Homogeneity*Income Inequality",
              "homclass3_V_gc:dclass3res_VIntermediate class (III+IV):gini_disp" = "Homogeneity*Intermediate Class*Income Inequality",
              "homclass3_V_gc:dclass3res_VWorking Class (V+VI+VII):gini_disp" = "Homogeneity*Working Class *Income Inequality")

\pagebreak

```{r tab2-ocde, results='asis', cache=TRUE}
set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1")
options(ggplot2.discrete.fill = set1)
options(ggplot2.discrete.colour = set1)
ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(text=ggplot2::element_text(size=15,  family="serif"))
options(scipen=999)
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal,texreg,MLMusingR,lme4)  
rm(list=ls())

dfreg <- df1 %>% dplyr::select(
  egal = egal2,
  # starts_with("class3"),
  # starts_with("digclass3"),
  # starts_with("dclass3"),
  # starts_with("homclass"),
  dclass3res_V,
  dclass3spo_V,
  homclass3_V_res,
  know_total,
  socialtrust,
  Q03pcm,
  edyears,
  female,
  agenum,
  religion,
  partner,
  union,
  workst,
  WEIGHT,
  region,
  "gini_disp"=wid_gini_disp,
  "gini_mkt",
  gv_spen,
  rel_red,
  abs_red,
  ilo_taxrev,
  ilo_govspe,
  d10d1=wid_p90p10,
  wid_p90p50,
  top10=wid_sharetop10,
  rgdpna,
  gdppercapita,
  oecd,
  country2, country, country3,
) %>%
  filter(oecd == "OECD") %>%
  filter(country2 != "SVN") %>% 
  na.omit() %>% 
    mutate(loggdppercapita=scale(gdppercapita/1000),
         edyears2 = edyears ^ 2,
         age2 = agenum ^ 2, 
         gini_disp=scale(gini_disp))

```{r fig5-ocde, fig.dim=c(12,7),fig.cap='Three-way interaccion effects for Redistributive Preferences, Network segregation, Social class and Income Inequality'}
### Predicted values homo*class*GiniD ---------------------------------------
# ginid_max<- round(max(dfreg$gini_disp) ,2)
# ginid_min<- round(min(dfreg$gini_disp),2)
